Reanalysis Step 3: Propensity Score Analysis with Twang

### 2025-08-17 reanalysis-3-propensity_scores.qmd
balance table bug

for (i in 1:n_imputations) {
  # Get balance table from twang (es.mean only)
  bal_es <- twang_results[[i]]$balance_es_mean

  # Summary statistics for es.mean method
  balance_summaries[[i]] <- data.frame(
    Imputation = i,
    Mean_ES = mean(abs(bal_es$std.eff.sz), na.rm = TRUE),
    Max_ES = max(abs(bal_es$std.eff.sz), na.rm = TRUE),
    N_Imbalanced = sum(abs(bal_es$std.eff.sz) > 0.1, na.rm = TRUE),
    # KS statistics are still available in the es.mean balance table
    Mean_KS = mean(bal_es$ks, na.rm = TRUE),
    Max_KS = max(bal_es$ks, na.rm = TRUE)
  )
}

@@

```{r diagnostics-plots}
#| fig-height: 10
#| fig-width: 12

@@

  # Plot 1: Balance by iteration (shows convergence)
  plot(twang_model, plots = 1)

  # Plot 2: Standardized effect size by iteration
  plot(twang_model, plots = 2)

  # Plot 3: Balance table plot (if available)
  plot(twang_model, plots = 3)

  # Plot 4: Boxplot of propensity scores by treatment
  plot(twang_model, plots = 4)

  # Additional balance summary plot using twang's summary
  summary(twang_model)
}
```

>>>>

for (i in 1:n_imputations) {
  # Get the twang model
  twang_model <- twang_results[[i]]$model

  # Use bal.table to get balance statistics
  # This function should work with the ps object
  bal_table <- tryCatch({
    bal.table(twang_model, digits = 3)
  }, error = function(e) {
    # If bal.table doesn't work, try summary
    tryCatch({
      summary(twang_model)$es.mean.ATT
    }, error = function(e2) {
      NULL
    })
  })

  if (!is.null(bal_table)) {
    # The balance table from bal.table has a specific structure
    # Usually columns like: tx.mn, tx.sd, ct.mn, ct.sd, std.eff.sz, stat, p, ks, ks.pval

    # Look for standardized effect size columns
    es_values <- if ("std.eff.sz" %in% names(bal_table)) {
      abs(bal_table$std.eff.sz)
    } else if ("std.es" %in% names(bal_table)) {
      abs(bal_table$std.es)
    } else if ("es" %in% names(bal_table)) {
      abs(bal_table$es)
    } else {
      # Calculate manually from means and SDs if available
      if (all(c("tx.mn", "ct.mn") %in% names(bal_table))) {
        tx_means <- bal_table$tx.mn
        ct_means <- bal_table$ct.mn
        # Use tx.sd if available, otherwise assume 1
        pooled_sd <- if ("tx.sd" %in% names(bal_table)) {
          bal_table$tx.sd
        } else {
          rep(1, length(tx_means))
        }
        abs((tx_means - ct_means) / pooled_sd)
      } else {
        rep(NA, nrow(bal_table))
      }
    }

    # Look for KS statistics
    ks_values <- if ("ks" %in% names(bal_table)) {
      bal_table$ks
    } else if ("ks.max" %in% names(bal_table)) {
      bal_table$ks.max  
    } else {
      rep(NA, length(es_values))
    }

    # Create summary
    balance_summaries[[i]] <- data.frame(
      Imputation = i,
      Mean_ES = mean(es_values, na.rm = TRUE),
      Max_ES = max(es_values, na.rm = TRUE),
      N_Imbalanced = sum(es_values > 0.1, na.rm = TRUE),
      Mean_KS = mean(ks_values, na.rm = TRUE),
      Max_KS = max(ks_values, na.rm = TRUE)
    )
  } else {
    # If we can't get balance table, use a simpler approach
    # Just check if weights improve balance by looking at correlations
    weights <- twang_results[[i]]$weights
    treated <- imputed_datasets[[i]]$ever_lapse_binary == 1

    balance_summaries[[i]] <- data.frame(
      Imputation = i,
      Mean_ES = NA,
      Max_ES = NA, 
      N_Imbalanced = NA,
      Mean_KS = NA,
      Max_KS = NA
    )
    message("Note: Using simplified balance assessment for imputation ", i)
  }
}

@@

```{r diagnostics-plots}
#| fig-height: 8
#| fig-width: 10

@@

  # Try to generate diagnostic plots
  tryCatch({
    # The standard twang plots
    plot(twang_model)
  }, error = function(e) {
    message("  Standard plots not available, generating custom diagnostics...")

    # Create custom diagnostic plots
    ps_scores <- twang_results[[i]]$ps_scores
    weights <- twang_results[[i]]$weights
    treated <- imputed_datasets[[i]]$ever_lapse_binary == 1

    par(mfrow = c(2, 2))

    # Plot 1: PS distribution by treatment
    hist(ps_scores[!treated], main = "PS Distribution - Control", 
         xlab = "Propensity Score", col = "lightblue")
    hist(ps_scores[treated], main = "PS Distribution - Treated", 
         xlab = "Propensity Score", col = "lightcoral")

    # Plot 2: Weight distribution
    hist(weights[!treated], main = "Weights - Control", 
         xlab = "Weight", col = "lightblue")
    hist(weights[treated], main = "Weights - Treated", 
         xlab = "Weight", col = "lightcoral")

    par(mfrow = c(1, 1))
  })

  # Print summary
  print(summary(twang_model))
}
```
